---
title: "Random Forest on Cleveland Heart Disease Data"
author: "PC Manefeldt"
date: "2024-12-05"
output: html_notebook
---


# Setup

```{r}
rm(list=ls())
gc()
```


```{r}
library('ranger')
library('tuneRanger')
library('vegan') # For stepacross function in ISOMAP functions
library(ggplot2); library(tibble) # Variable Importance Plot
library('mapview'); library("leaflet"); library("leafpop") # Interactive plotting
source("./IsomapFunctions.R")
source("./ThesisScript.R")
source("./PCO_ACP_Plotting.R")
```


# Data

```{r}
heart.dat <- read.csv('./heart+disease/processed.cleveland.data',header = FALSE)
heart.dat <- na.omit(heart.dat)
heart.dat <- heart.dat[-c(which(sapply(1:nrow(heart.dat),function(x)any(heart.dat[x,]=="?")))),]
colnames(heart.dat) <- c('age', 'sex', 'cp', 'trestbps', 'chol','fbs','restecg','thalach','exang','oldpeak','slope','ca','thal','num')
heart.df <- data.frame('age'=heart.dat$age, 
                       'sex'=factor(heart.dat$sex),
                       'cp'=factor(heart.dat$cp),
                       'trestbps'=heart.dat$trestbps,
                       'chol'=heart.dat$chol,
                       'fbs'=factor(heart.dat$fbs),
                       'restecg'=factor(heart.dat$restecg),
                       'thalach'=heart.dat$thalach,
                       'exang'=factor(heart.dat$exang),
                       'oldpeak'=heart.dat$oldpeak,
                       'slope'=factor(heart.dat$slope),
                       'ca'=factor(heart.dat$ca,ordered = TRUE),
                       'thal'=factor(heart.dat$thal),
                       'Class'=(heart.dat$num))
heart.df$Class <- factor(ifelse(heart.df$Class==0,0,
                         ifelse(heart.df$Class==1,1,2)))
```


# Random Forest

## Tuning

```{r}
set.seed(123)
heart.task <- makeClassifTask(data=heart.df, target="Class")

estimateTimeTuneRanger(heart.task)
# Tuning
res = tuneRanger(heart.task, num.trees = 1000,measure = list(multiclass.brier),
  num.threads = 4, iters = 70,iters.warmup=50, save.file.path = NULL, 
  parameters=list(sample.fraction=1,replace=TRUE),
  tune.parameters=c("mtry", "min.node.size"))
# Mean of best 5 % of the results
res
# Model with the new tuned hyperparameters
res$model
tmp.mod <- ranger(Class~.,data=heart.df[,], importance="permutation",
                  mtry=res$recommended.pars$mtry,
                  min.node.size = res$recommended.pars$min.node.size, 
                  sample.fraction =1,
                  num.trees = 2000,class.weights = 1/table(heart.df$Class)/sum(1/table(heart.df$Class)),
                  keep.inbag=TRUE)
```


## Training

```{r}
set.seed(123)
heart.mod <- tmp.mod <- mod <- ranger(Class~.,data=heart.df[,], importance="permutation",
                  mtry=3,
                  min.node.size = 21, 
                  sample.fraction = 1,
                  class.weights = 1/table(heart.df$Class)/sum(1/table(heart.df$Class)),
                  num.trees = 2000,
                  keep.inbag=TRUE)
tmp.dat <- heart.df
```

# Performance

## OOB performance

```{r}
tmp.mod$confusion.matrix
tmp.mod$prediction.error
```

```{r}
knitr::kable(tmp.mod$confusion.matrix,format = 'latex')
1-tmp.mod$prediction.error
```


## Training performance

```{r}
hh <- predict(tmp.mod,heart.df)
table(hh$predictions,heart.df$Class)
sum(diag(table(hh$predictions,heart.df$Class)))/nrow(heart.df)
```

# Variable Imporatance Plot

```{r}
#pdf("plots/heartvarimp.pdf")
ggplot(
    enframe(
        tmp.mod$variable.importance,
        name = "variable",
        value = "importance"
    ),
    aes(
        x = reorder(variable, importance),
        y = importance,
        fill = importance
    )
) +
    geom_bar(stat = "identity", position = "dodge") +
    coord_flip() +
    ylab("Variable Importance") +
    xlab("") +
    #ggtitle("Permutation Variable Importance Plot") +
    guides(fill = "none") +
    scale_fill_gradient(low = "red", high = "blue")+ 
    theme_classic()
#dev.off()
```


# Original/Full Proximity Measure

```{r, fig.asp=1}
aa <- extract_proximity_full(fit=mod,olddata = tmp.dat)
#aa <- extract_proximity_oob(fit=mod,olddata = tmp.dat)
dissim.mat1 <- sqrt(1-aa)
```





# Proximity Plot with added variable information

```{r}
eig.isomap.sol <- eig.isomap.fun(dist.mat = dissim.mat1,eps=1.1,#eps = eps.opt.list$opt.eps,
                                 Trace=FALSE, shortest=FALSE)
tmp.coords.orig <- tmp.coords <- eigs.to.coords.fun(eig.isomap.sol$eig.obj,p=2)[[1]][,1:2]
mds.obj.tmp <- list(coords=tmp.coords)
tmp.coords <- rot.plot.fun(mds.obj.tmp,rot.angle=-30)[[1]]
coords.cad.prox <- tmp.coords
```

```{r}
tmp.dist <- as.matrix(dist(tmp.coords[,1:2]))
tmp.dist <- as.matrix(dist(tmp.coords.orig[,1:2]))
1-sum(sqrt((eig.isomap.sol$dissim.mat^2-tmp.dist^2)^2))/sum(eig.isomap.sol$dissim.mat^2)
#tmp.dist[1:5,1:5]
cumsum(eig.isomap.sol$eig.obj$values[1:100])/sum((eig.isomap.sol$eig.obj$values)[eig.isomap.sol$eig.obj$values>0])
```

```{r}
Class <- heart.df$Class
Class.int <- as.numeric(Class)
class.vec <- adjustcolor(RColorBrewer::brewer.pal(n=length(unique(Class)), name='YlOrRd'),alpha.f = 0.9)
class.cols <- class.vec[Class.int]

age.int <- cut(heart.df$age,breaks=8)
age.vec <- adjustcolor(RColorBrewer::brewer.pal(n=length(unique(age.int)), name='YlOrRd'),alpha.f = 1)
age.cols <- age.vec[as.factor(age.int)]

Ca.int <- as.numeric(heart.df$ca)-1
ca.vec <- adjustcolor(RColorBrewer::brewer.pal(n=length(unique(Ca.int)), name='YlOrRd'),alpha.f = 1)
ca.cols <- ca.vec[as.factor(Ca.int)]

oldpeak.int <- cut(as.numeric(heart.df$oldpeak),breaks=c(0,0.5,1:7),include.lowest = TRUE,right=FALSE)
oldpeak.vec <- adjustcolor(RColorBrewer::brewer.pal(n=length(unique(oldpeak.int)), name='YlOrRd'),alpha.f = 1)
oldpeak.cols <- oldpeak.vec[as.factor(oldpeak.int)]


thal.int <- ifelse(heart.df$thal=="3.0",1,ifelse(heart.df$thal=="6.0",2,3))
cp.int <- as.numeric(heart.df$cp)
slope.int <- as.numeric(heart.df$slope)
```

## Class

```{r,fig.asp=1}
#pdf("plots/heartproxplot.pdf")
par(mar=c(2.1, 2.1, 0.1, 0.1))
plot(tmp.coords.orig,
     asp=1,
     col=class.cols,
     pch=16,
     xaxt='n',yaxt='n', ylab="",xlab="", cex=1)
mtext("Dim 1",side=1,line = 0.5,cex = 1.2)
mtext("Dim 2",side=2,line = 0.5,cex = 1.2)
legend(x='topright',col=class.vec,title="CAD Diagnosis",
       legend = c("No CAD","CAD","MCAD"),
       pch=16)
#dev.off()
```


## Sex and class

```{r,fig.asp=1}
#pdf("plots/heartproxplotsex.pdf")
par(mar=c(2.1, 2.1, 0.1, 0.1))
plot(tmp.coords,
     asp=1,
     col=class.cols,
     pch=c(16,17)[heart.df$sex],
     xaxt='n',yaxt='n', ylab="",xlab="", cex=1)
mtext("Dim 1",side=1,line = 0.5,cex = 1.2)
mtext("Dim 2",side=2,line = 0.5,cex = 1.2)
legend(x='topright',col=class.vec,
       legend = c("No CAD","CAD","MCAD"),
       pch=16,title="CAD Diagnosis")
legend(x = "bottomright",col="black",
       legend = c("Female","Male"),
       title="Sex",
       pch=c(16,17))
#dev.off()
```

## Thallium Stress Scintigraphy and Fluoroscopy

```{r,fig.asp=1}
#pdf("plots/heartproxplotca.pdf")
par(mar=c(2.1, 2.1, 0.1, 0.1))
plot(tmp.coords,
     asp=1,
     col=ca.cols,
     pch=thal.int+14,#c(15,16,17,18,19,20)[as.numeric(heart.df$ca)],
     xaxt='n',yaxt='n', ylab="",xlab="", cex=1)
mtext("Dim 1",side=1,line = 0.5,cex = 1.2)
mtext("Dim 2",side=2,line = 0.5,cex = 1.2)
legend(x='topright',col=ca.vec,
       legend = c(0:3),#c(sort(unique(Ca.int))-1),#seq(0,length(unique(heart.df$Class))-1,1),
       pch=16,title="Number of Major vessels \n coloured by fluoroscopy")
legend(x = "bottom",col="black",
       legend = c("Normal","Fixed Defect","Reversible Defect"),#seq(0,length(unique(heart.df$Class))-1,1),
       pch=c(15,16,17),title="Thallium Stress \n Scintigraphy")
#dev.off()
```

## Chest Pain and Age

```{r,fig.asp=1}
#pdf("plots/heartproxplotcp.pdf")
par(mar=c(2.1, 2.1, 0.1, 0.1))

set.seed(123)
plot(tmp.coords,
     asp=1,
     col=age.cols,
     pch=as.numeric(cp.int)+14,
     xaxt='n',yaxt='n', ylab="",xlab="", cex=1)
mtext("Dim 1",side=1,line = 0.5,cex = 1.2)
mtext("Dim 2",side=2,line = 0.5,cex = 1.2)
legend(x='bottomright',col=age.vec,
       legend = levels(age.int),
       pch=16,title="Patient Age")
legend(x = "bottom",col="black",
       legend = c("typical angina", "atypical angina", "non-anginal pain","asymptomatic"),
       pch=c(15,16,17,18),title="Type of Chest Pain")
#dev.off()
```


## ECG ST wave measures (slope and oldpeak)

```{r,fig.asp=1}
#pdf("plots/heartproxplotoldpeak.pdf")
par(mar=c(2.1, 2.1, 0.1, 0.1))

set.seed(123)
plot(tmp.coords,
     asp=1,
     col=oldpeak.cols,
     pch=as.numeric(slope.int)+14,
     xaxt='n',yaxt='n', ylab="",xlab="", cex=1)
mtext("Dim 1",side=1,line = 0.5,cex = 1.2)
mtext("Dim 2",side=2,line = 0.5,cex = 1.2)
legend(x='bottomright',col=age.vec,
       legend = levels(oldpeak.int),
       pch=16,title="ST depression (oldpeak)")
legend(x = "bottom",col="black",
       legend = c("Upsloping", "Flat", "Downsloping"),
       pch=c(15,16,17),title="ST Segment (slope)")
#dev.off()
```


## Blank (was used for regional interpretation)

```{r,fig.asp=1}
#pdf("plots/heartproxplotblank.pdf")

par(mar=c(2.1, 2.1, 0.1, 0.1))
plot(tmp.coords,
     asp=1,
     col=class.cols,
     pch=16,#c(16:17)[as.numeric(heart.df$sex)],
     xaxt='n',yaxt='n', ylab="",xlab="", cex=1)
mtext("Dim 1",side=1,line = 0.5,cex = 1.2)
mtext("Dim 2",side=2,line = 0.5,cex = 1.2)
legend(x='topright',col=class.vec,
       legend = c("No CAD","CAD","MCAD"),
       pch=16)
legend(x = 0.752,y=0.85,col="black",
       legend = c("Female","Male"),
       pch=c(16,17))
#dev.off()
```


# ISOMAP 

```{r}
eig.isomap.sol <- eig.isomap.fun(dist.mat = dissim.mat1,eps=1,#eps = eps.opt.list$opt.eps,
                                 Trace=FALSE, shortest=FALSE)
tmp.coords.orig<- tmp.coords <- eigs.to.coords.fun(eig.isomap.sol$eig.obj,p=2)[[1]][,1:2]
tmp.coords <- ortho.procrustes(Target = coords.cad.prox,Testee=tmp.coords)[[1]]
ortho.procrustes(Target = coords.cad.prox,Testee=tmp.coords)[[2]]
```

```{r}
#tmp.dist <- as.matrix(dist(tmp.coords[,1:2]))
tmp.dist <- as.matrix(dist(tmp.coords.orig[,1:2]))
1-sum(sqrt((eig.isomap.sol$dissim.mat^2-tmp.dist^2)^2))/sum(eig.isomap.sol$dissim.mat^2)
#tmp.dist[1:5,1:5]
cumsum(eig.isomap.sol$eig.obj$values[1:100])/sum((eig.isomap.sol$eig.obj$values)[eig.isomap.sol$eig.obj$values>0])
```

```{r,fig.asp=1}
#pdf("plots/heartisomapplot.pdf")
par(mar=c(2.1, 2.1, 0.1, 0.1))

plot(tmp.coords,
     asp=1,
     col=class.cols,
     pch=16,
     xaxt='n',yaxt='n', ylab="",xlab="", cex=1)
mtext("Dim 1",side=1,line = 0.5,cex = 1.2)
mtext("Dim 2",side=2,line = 0.5,cex = 1.2)
legend(x='topright',col=class.vec,title="CAD Diagnosis",
       legend = c("No CAD","CAD","MCAD"),
       pch=16)
#dev.off()
```

## Sex and Class

```{r,fig.asp=1}
#pdf("plots/heartisomapplotsex.pdf")
par(mar=c(2.1, 2.1, 0.1, 0.1))
plot(tmp.coords,
     asp=1,
     col=class.cols,
     pch=c(16,17)[heart.df$sex],
     xaxt='n',yaxt='n', ylab="",xlab="", cex=1)
mtext("Dim 1",side=1,line = 0.5,cex = 1.2)
mtext("Dim 2",side=2,line = 0.5,cex = 1.2)
legend(x='topright',col=class.vec,
       legend = c("No CAD","CAD","MCAD"),
       pch=16,title="CAD Diagnosis")
legend(x = "bottomright",col="black",
       legend = c("Female","Male"),
       title="Sex",
       pch=c(16,17))
#dev.off()
```

## Thallium Stress Scintigraphy and Fluoroscopy

```{r,fig.asp=1}
#pdf("plots/heartisomapplotca.pdf")
par(mar=c(2.1, 2.1, 0.1, 0.1))
set.seed(123)
plot(tmp.coords,
     asp=1,
     col=ca.cols,
     pch=thal.int+14,
     xaxt='n',yaxt='n', ylab="",xlab="", cex=1)
mtext("Dim 1",side=1,line = 0.5,cex = 1.2)
mtext("Dim 2",side=2,line = 0.5,cex = 1.2)
legend(x='topright',col=ca.vec,
       legend = c(0:3),#c(sort(unique(Ca.int))-1),#seq(0,length(unique(heart.df$Class))-1,1),
       pch=16,title="Number of Major vessels \n coloured by fluoroscopy")
legend(x = "bottom",col="black",
       legend = c("Normal","Fixed Defect","Reversible Defect"),#seq(0,length(unique(heart.df$Class))-1,1),
       pch=c(15,16,17),title="Thallium Stress \n Scintigraphy")
#dev.off()
```

## Chest Pain and Age

```{r,fig.asp=1}
#pdf("plots/heartisomapplotcp.pdf")
par(mar=c(2.1, 2.1, 0.1, 0.1))

set.seed(123)
plot(tmp.coords,
     asp=1,
     col=age.cols,
     pch=as.numeric(cp.int)+14,
     xaxt='n',yaxt='n', ylab="",xlab="", cex=1)
mtext("Dim 1",side=1,line = 0.5,cex = 1.2)
mtext("Dim 2",side=2,line = 0.5,cex = 1.2)
legend(x='bottomright',col=age.vec,
       legend = levels(age.int),
       pch=16,title="Patient Age")
legend(x = "bottom",col="black",
       legend = c("typical angina", "atypical angina", "non-anginal pain","asymptomatic"),
       pch=c(15,16,17,18),title="Type of Chest Pain")
#dev.off()
```


## ECG ST wave measures (slope and oldpeak)

```{r,fig.asp=1}
#pdf("plots/heartisomapplotoldpeak.pdf")
par(mar=c(2.1, 2.1, 0.1, 0.1))

set.seed(123)
plot(tmp.coords,
     asp=1,
     col=oldpeak.cols,
     pch=as.numeric(slope.int)+14,
     xaxt='n',yaxt='n', ylab="",xlab="", cex=1)
mtext("Dim 1",side=1,line = 0.5,cex = 1.2)
mtext("Dim 2",side=2,line = 0.5,cex = 1.2)
legend(x='bottomright',col=age.vec,
       legend = levels(oldpeak.int),
       pch=16,title="ST depression (oldpeak)")
legend(x = "bottom",col="black",
       legend = c("Upsloping", "Flat", "Downsloping"),
       pch=c(15,16,17),title="ST Segment (slope)")
#dev.off()
```

# Bootstrap confidence regions

## PCO bootstrap confidence regions based on Tukey depth and alpha bags

```{r}
tmp.dat <- dat <- heart.df
tmp.mod <- heart.mod
Class.int <- as.numeric(heart.df$Class)
col.vec <- adjustcolor(RColorBrewer::brewer.pal(n=length(levels(Class)), name='YlOrRd'),alpha.f = 1)
cols <- col.vec[Class.int]

n <- nrow(dat)
set.seed(123)
p <- 2
B <- 100
boot.mat <- matrix(sample(1:n,size=round(n*B,digits = 0),replace=TRUE),nrow=n,ncol=B)
dissim.mat <- sqrt(1-extract_proximity_full(tmp.mod,dat))

  A.mat <- -0.5*dissim.mat^2
  col.centers <- colMeans(A.mat)
  row.centers <- colMeans(t(A.mat))
  center <-  mean(A.mat)
  centred_mat <- apply(apply(A.mat,1,function(x)x-col.centers),
                       1,function(x)x-row.centers)+center
  eig.dat <- eigen(centred_mat)
  coords <- eig.dat$vectors[,1:p,drop=FALSE]%*%diag(eig.dat$values[1:p],nrow=p)^0.5

```


```{r, fig.asp=1}
coords <- tmp.coords.orig
n <- length(coords)
boot.mds.map.list <- list()
proc.stat <- numeric()
for(i in 1:B){
  n <- nrow(tmp.dat)
  index <- boot.mat[,i]
  dist.mat <- dissim.mat[index,index]
  #dist.mat <- dissim.mat[index,]
  
  A.mat2 <- -0.5*dist.mat^2
  col.centers2 <- colMeans(A.mat2)
  row.centers2 <- colMeans(t(A.mat2))
  center2 <-  mean(A.mat2)
  centred_mat2 <- apply(apply(A.mat2,1,function(x)x-col.centers2),
                       1,function(x)x-row.centers2)+center2
  
  
  
  eig.dat2 <- eigen(centred_mat2)

  
  new.dissim <- dissim.mat
  new.A.mat <- -0.5*new.dissim^2
  new.row.centers <- colMeans(t(new.A.mat))
  new.row.centers[index] <- row.centers2
  new.col.centers <- colMeans(new.A.mat)
  new.col.centers[index] <- col.centers2
  #new.centred_mat <- apply(apply(new.A.mat,1,function(x)x-col.centers2),
  #                     1,function(x)x-new.row.centers)+center2
  new.centred_mat <- apply(apply(new.A.mat,1,function(x)x-new.col.centers),
                       1,function(x)x-new.row.centers)+center2
  
  new.eigen.vecs <- matrix(0,nrow=n,ncol=n)
  for(j in unique(index)){
    new.eigen.vecs[j,] <- apply(eig.dat2$vectors[which(index==j),,drop=FALSE],2,sum)
  }
  new.coords <- new.centred_mat%*%
    (new.eigen.vecs)%*%diag(eig.dat2$values^(-0.5))

  # Perform Orthogonal Procrustes
  fit.X <- ortho.procrustes(Target=coords[,1:2],Testee = new.coords[,1:2])
  coords2 <- fit.X[[1]]
  
  boot.mds.map.list[[i]] <- coords2
  proc.stat[i] <- fit.X[[2]]
}

boot.mds.map.array <- simplify2array(boot.mds.map.list)

```



```{r, fig.asp=1}
#pdf("plots/heartdiseasecRFalphabags.pdf")
a <- par()
par(mar=c(0.1,0.1,0.1,0.1))
alpha <- 0.90
grid.n <- 70

# Plotting spanning ellipses instead of circles
plot(
coords[,1:2],
asp=1,
col=cols,
main="",#main = paste("Classical Scaling of ", k, " Nearest Neighbours similarity \n with ", alpha*100,"% Bootstrap Confidence Regions",sep="") ,
xlab="",#paste("PCoA 1 (",round(eig.dat$values[1]/sum(eig.dat$values)*100,2),"%)",sep=""),
ylab="",#paste("PCoA 2 (",round(eig.dat$values[2]/sum(eig.dat$values)*100,2),"%)",sep=""),
pch=16, axes =FALSE, frame.plot=TRUE, cex=0.2
,xlim=range(coords[,1])+c(-0.11,0.05),ylim=range(coords[,2])+c(-0.05,0.1)
,type='n'
)

#mtext(paste("PCoA 1 (",round(eig.dat$values[1]/sum(eig.dat$values)*100,2),"%)",sep=""), side = 1,line = 1)
#mtext(paste("PCoA 2 (",round(eig.dat$values[2]/sum(eig.dat$values)*100,2),"%)",sep=""), side = 2,line = 1)

legend(x = "bottomright",legend=levels(heart.df$Class), lty=1, col=col.vec)

#cols <- rep(c(rgb(0,0,0,0.5),rgb(1,0,0,0.5),rgb(0,1,0,0.5)),each=50)
a <- sapply(1:dim(boot.mds.map.array)[1],function(x){
  boot.tmp <- t(boot.mds.map.array[x,,])
  boot.tukey.depth <- ddalpha::depth.halfspace(x=boot.tmp,data=boot.tmp, num.directions = 200)
  grid.points <- as.matrix(expand.grid(seq(from=min(boot.tmp[,1]),to=max(boot.tmp[,1]),length.out=grid.n),
                           seq(from=min(boot.tmp[,2]),to=max(boot.tmp[,2]),length.out=grid.n)))
  grid.points.tukey.depth <- ddalpha::depth.halfspace(x=grid.points,data=boot.tmp, num.directions = 200)
  grid.include.ellipse <-which(grid.points.tukey.depth>=quantile(boot.tukey.depth,1-alpha))

  Plot_ConvexHull(xcoord=grid.points[grid.include.ellipse,1],ycoord=grid.points[grid.include.ellipse,2], lcolor=cols[x])
})

#dev.off()
```

## Isomap bootstrap confidence regions based on Tukey depth and alpha bags

```{r}
set.seed(123)
library('vegan')
tmp.dat <- dat <- heart.df
tmp.mod <- heart.mod
Class.int <- as.numeric(heart.df$Class)
col.vec <- adjustcolor(RColorBrewer::brewer.pal(n=length(levels(Class)), name='YlOrRd'),alpha.f = 1)
cols <- col.vec[Class.int]

n <- nrow(dat)
#set.seed(123)
B <- 100
boot.mat <- matrix(sample(1:n,size=round(n*B,digits = 0),replace=TRUE),nrow=n,ncol=B)
dissim.mat.org <- sqrt(1-extract_proximity_full(tmp.mod,dat))
    s.paths <- stepacross(dissim.mat.org,
                            path='extended',
                            toolong=1, 
                            trace = FALSE)

    dissim.mat <- as.matrix(s.paths)

  A.mat <- -0.5*dissim.mat^2
  col.centers <- colMeans(A.mat)
  row.centers <- colMeans(t(A.mat))
  center <-  mean(A.mat)
  centred_mat <- apply(apply(A.mat,1,function(x)x-col.centers),
                       1,function(x)x-row.centers)+center
  eig.dat <- eigen(centred_mat)
  coords <- eig.dat$vectors[,1:p,drop=FALSE]%*%diag(eig.dat$values[1:p],nrow=p)^0.5

```


```{r, fig.asp=1}
boot.mds.map.list <- list()
proc.stat <- numeric()
for(i in 1:B){
  n <- nrow(tmp.dat)
  index <- boot.mat[,i]
  dist.mat <- dissim.mat.org[index,index]
  s.paths <- stepacross(dist.mat,
                        path='extended',
                        toolong=1, 
                        trace = FALSE)

  dist.mat <- as.matrix(s.paths)
  
  A.mat2 <- -0.5*dist.mat^2
  col.centers2 <- colMeans(A.mat2)
  row.centers2 <- colMeans(t(A.mat2))
  center2 <-  mean(A.mat2)
  centred_mat2 <- apply(apply(A.mat2,1,function(x)x-col.centers2),
                       1,function(x)x-row.centers2)+center2
  
  
  
  eig.dat2 <- eigen(centred_mat2)

  new.dissim <- dissim.mat
  #new.dissim <- dissim.mat.org
  #new.dissim[index,index] <- dist.mat
  #new.dissim[-index,-index] <-   #ifelse(new.dissim[-index,-index]>=1,NA,new.dissim[-index,-index])
  
  #s.paths <- stepacross(new.dissim,
  #                      path='extended',
  #                      toolong=max(new.dissim)+0.1, 
  #                      trace = FALSE)
#  new.dissim <- as.matrix(s.paths)
  new.A.mat <- -0.5*new.dissim^2
  new.row.centers <- colMeans(t(new.A.mat))
  new.row.centers[index] <- row.centers2
  new.col.centers <- colMeans(new.A.mat)
  new.col.centers[index] <- col.centers2
  #new.centred_mat <- apply(apply(new.A.mat,1,function(x)x-col.centers2),
  #                     1,function(x)x-new.row.centers)+center2
  new.centred_mat <- apply(apply(new.A.mat,1,function(x)x-new.col.centers),
                       1,function(x)x-new.row.centers)+center2
  
  new.eigen.vecs <- matrix(0,nrow=n,ncol=n)
  for(j in unique(index)){
    new.eigen.vecs[j,] <- apply(eig.dat2$vectors[which(index==j),,drop=FALSE],2,sum)
  }
  new.coords <- new.centred_mat%*%
    (new.eigen.vecs)%*%diag(eig.dat2$values^(-0.5))

  # Perform Orthogonal Procrustes
  fit.X <- ortho.procrustes(Target=coords[,1:2],Testee = new.coords[,1:2])
  coords2 <- fit.X[[1]]
  
  boot.mds.map.list[[i]] <- coords2
  proc.stat[i] <- fit.X[[2]]
}

boot.mds.map.array <- simplify2array(boot.mds.map.list)

```



```{r, fig.asp=1}

#pdf("plots/heartdiseaseIsoRFalphabags.pdf")
par(mar=c(0.1,0.1,0.1,0.1))
alpha <- 0.90
grid.n <- 70

# Plotting spanning ellipses instead of circles
plot(
coords[,1:2],
asp=1,
col=cols,
main="",
xlab="",
ylab="",
pch=16, axes =FALSE, frame.plot=TRUE, cex=1
,xlim=range(coords[,1])+c(-0.05,0.05),ylim=range(coords[,2])+c(-0.05,0.05)
,type='n'
)

#mtext(paste("PCoA 1 (",round(eig.dat$values[1]/sum(eig.dat$values)*100,2),"%)",sep=""), side = 1,line = 1)
#mtext(paste("PCoA 2 (",round(eig.dat$values[2]/sum(eig.dat$values)*100,2),"%)",sep=""), side = 2,line = 1)

legend(x = "bottomright",legend=levels(heart.df$Class), lty=1, col=col.vec)

#cols <- rep(c(rgb(0,0,0,0.5),rgb(1,0,0,0.5),rgb(0,1,0,0.5)),each=50)
a <- sapply(1:dim(boot.mds.map.array)[1],function(x){
  boot.tmp <- t(boot.mds.map.array[x,,])
  boot.tukey.depth <- ddalpha::depth.halfspace(x=boot.tmp,data=boot.tmp, num.directions = 200)
  grid.points <- as.matrix(expand.grid(seq(from=min(boot.tmp[,1]),to=max(boot.tmp[,1]),length.out=grid.n),
                           seq(from=min(boot.tmp[,2]),to=max(boot.tmp[,2]),length.out=grid.n)))
  grid.points.tukey.depth <- ddalpha::depth.halfspace(x=grid.points,data=boot.tmp, num.directions = 200)
  grid.include.ellipse <-which(grid.points.tukey.depth>=quantile(boot.tukey.depth,1-alpha))

  Plot_ConvexHull(xcoord=grid.points[grid.include.ellipse,1],ycoord=grid.points[grid.include.ellipse,2], lcolor=cols[x])
})

#dev.off()
```


# Interactive Plotting via Mapview

```{r}
tmp.terra <- terra::vect(tmp.coords)
preds <- predict(tmp.mod,heart.df)$predictions
oob.preds <- tmp.mod$predictions
tmp.imp.info <- cbind(heart.df,preds,oob.preds)
tmp.imp.info$thal <- ifelse(tmp.imp.info$thal=="3.0","Normal",ifelse(tmp.imp.info$thal=="6.0","Fixed Defect","Reversable Defect"))
tmp.imp.info$Class <- ifelse(tmp.imp.info$Class=="0","No CAD",
                             ifelse(tmp.imp.info$Class=="1","CAD",
                                    "MCAD"))
tmp.terra <- cbind(tmp.terra,tmp.imp.info)


Heart <- tmp.terra
mapview::mapview(Heart,zcol=c("Class","preds","oob.preds","cp","sex","age","thal","ca","oldpeak","thalach","restecg"),popup = popupTable(heart.df),
                 layer.name=c("True Diagnosis",
                              "Predicted Diagnosis",
                              "OOB Predicted Diagnosis",
                              "Chest Pain",
                              "Sex",
                              "Age",
                              "Thallium Stress Scintigraph",
                              "Number of Major Vessels in Fluoroscopy",
                              "oldpeak","thalach","restecg"))

```


