---
title: "Random Forest on Palmer Penguins Data"
author: "PC Manefeldt"
date: "2024-12-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Setup

```{r}
rm(list=ls())
gc()
```


```{r}
library('ranger')
library('tuneRanger')
library('vegan') # For stepacross function in ISOMAP functions
library(ggplot2); library(tibble) # Variable Importance Plot
library('mapview'); library("leaflet"); library("leafpop") # Interactive plotting
source("./IsomapFunctions.R")
source("./ThesisScript.R")
source("./PCO_ACP_Plotting.R")
```


# Data

```{r}
source("../Palmer Penguins Example/PenguinsData.R")
class.labels <- as.factor(as.matrix(dat[,1])[,1])
predictor.mat <- as.matrix(dat[,-1])
```

# Random Forest

## Tuning

```{r}
penguin.task <- makeClassifTask(data=dat, target="species")
estimateTimeTuneRanger(penguin.task)
# Tuning
res = tuneRanger(penguin.task, measure = list(multiclass.brier), num.trees = 2000,parameters = list(replace = TRUE),num.threads = 4, iters = 100, iters.warmup = 100, save.file.path = NULL)
# Mean of best 5 % of the results
res
#Model with the new tuned hyperparameters
res$model
```

## Training

```{r}
set.seed(123)
penguin.mod <- tmp.mod <- ranger(species~.,data=dat, importance="permutation",
                  mtry=2,
                  min.node.size = 2, 
                  sample.fraction =0.8264299,
                  num.trees = 2000,
                  keep.inbag=TRUE)
```
# Performance

## OOB performance

```{r}
tmp.mod$confusion.matrix
tmp.mod$prediction.error
```

```{r}
knitr::kable(tmp.mod$confusion.matrix,format = 'latex')
1-tmp.mod$prediction.error
```


## Training performance

```{r}
hh <- predict(tmp.mod,dat)
table(hh$predictions,dat$species)
sum(diag(table(hh$predictions,dat$species)))/nrow(dat)
```


```{r}
#pdf("plots/heartvarimp.pdf")
ggplot(
    enframe(
        tmp.mod$variable.importance,
        name = "variable",
        value = "importance"
    ),
    aes(
        x = reorder(variable, importance),
        y = importance,
        fill = importance
    )
) +
    geom_bar(stat = "identity", position = "dodge") +
    coord_flip() +
    ylab("Variable Importance") +
    xlab("") +
    #ggtitle("Permutation Variable Importance Plot") +
    guides(fill = "none") +
    scale_fill_gradient(low = "red", high = "blue")+ 
    theme_classic()
#dev.off()
```

# Original/Full Proximity Measure

```{r, fig.asp=1}
tmp.dat <- dat
mod <- tmp.mod
aa <- extract_proximity_full(fit=mod,olddata = tmp.dat)
dissim.mat1 <- sqrt(1-aa)
```

# Plotting


```{r}
tmp.dat <- dat[,-1]#[1:200,]
Class <- dat$species


Class.int <- as.numeric(Class)
#col.vec <- c(rgb(0,0,0,0.8),rgb(1,0,0,0.8),rgb(0,1,0,0.8))
col.vec <- adjustcolor(RColorBrewer::brewer.pal(n=length(levels(Class)), name='Dark2'),alpha.f = 0.8)
#col.vec <- adjustcolor(RColorBrewer::brewer.pal(n=length(levels(Class)), name='YlOrRd'),alpha.f = 0.5)
#col.vec <- adjustcolor(RColorBrewer::brewer.pal(n=length(levels(Class)), name='YlOrRd'),alpha.f = 1)
cols <- col.vec[Class.int]
mod <- tmp.mod
```


## PCO Proximity plot with classes added

```{r,fig.asp=1}
#pdf("plots/palmerproxplot.pdf")
par(mar=c(2.1, 2.1, 0.1, 0.1))
# Epsilon of 0.88 gives 1 dimensional solution
eig.isomap.sol <- eig.isomap.fun(dist.mat = dissim.mat1[], eps=1.1,
                                 Trace=FALSE, shortest=FALSE)
tmp.coords <- eigs.to.coords.fun(eig.isomap.sol$eig.obj,p=2)[[1]][,1:2]
plot(tmp.coords,
     asp=1,
     col=cols,
     pch=16,,
     xaxt='n',yaxt='n', ylab="",xlab="", cex=1)
mtext("Dim 1",side=1,line = 0.5,cex = 1.2)
mtext("Dim 2",side=2,line = 0.5,cex = 1.2)
legend(x="bottomright",col=col.vec,
       legend = levels(Class),
       pch=16)
#dev.off()
```


## Isomap proximity plot with class

```{r,fig.asp=1}
#pdf("plots/palmerisomapplot.pdf")
par(mar=c(2.1, 2.1, 0.1, 0.1))
# Epsilon of 0.88 gives 1 dimensional solution
eig.isomap.sol <- eig.isomap.fun(dist.mat = dissim.mat1[], eps=1,
                                 Trace=FALSE, shortest=FALSE)
tmp.coords <- eigs.to.coords.fun(eig.isomap.sol$eig.obj,p=2)[[1]][,1:2]
plot(tmp.coords,
     asp=1,
     col=cols,
     pch=16,#c(16,17)[heart.df$sex],
     xaxt='n',yaxt='n', ylab="",xlab="", cex=1)
mtext("Dim 1",side=1,line = 0.5,cex = 1.2)
mtext("Dim 2",side=2,line = 0.5,cex = 1.2)
legend(x="bottomright",col=col.vec,
       legend = levels(Class),
       pch=16)
#dev.off()

```



# Bootstrap confidence regions

## PCO bootstrap confidence regions based on Tukey depth and alpha bags
```{r}
tmp.mod <- penguin.mod

n <- nrow(dat)
set.seed(123)
B <- 100
p <- 2
boot.mat <- matrix(sample(1:n,size=round(n*B,digits = 0),replace=TRUE),nrow=n,ncol=B)
dissim.mat <- sqrt(1-extract_proximity_full(tmp.mod,dat))

  A.mat <- -0.5*dissim.mat^2
  col.centers <- colMeans(A.mat)
  row.centers <- colMeans(t(A.mat))
  center <-  mean(A.mat)
  centred_mat <- apply(apply(A.mat,1,function(x)x-col.centers),
                       1,function(x)x-row.centers)+center
  eig.dat <- eigen(centred_mat)
  coords <- eig.dat$vectors[,1:p,drop=FALSE]%*%diag(eig.dat$values[1:p],nrow=p)^0.5
```



```{r, fig.asp=1}
boot.mds.map.list <- list()
proc.stat <- numeric()
for(i in 1:B){
  n <- nrow(tmp.dat)
  index <- boot.mat[,i]
  dist.mat <- dissim.mat[index,index]
  
  A.mat2 <- -0.5*dist.mat^2
  col.centers2 <- colMeans(A.mat2)
  row.centers2 <- colMeans(t(A.mat2))
  center2 <-  mean(A.mat2)
  centred_mat2 <- apply(apply(A.mat2,1,function(x)x-col.centers2),
                       1,function(x)x-row.centers2)+center2
  
  
  
  eig.dat2 <- eigen(centred_mat2)
  # The coordinates for the factual observations
  coords2 <- eig.dat2$vectors
  
  
  new.dissim <- dissim.mat
  new.A.mat <- -0.5*new.dissim^2
  new.row.centers <- colMeans(t(new.A.mat))
  new.row.centers[index] <- row.centers2
  new.col.centers <- colMeans(new.A.mat)
  new.col.centers[index] <- col.centers2
  new.centred_mat <- apply(apply(new.A.mat,1,function(x)x-new.col.centers),
                       1,function(x)x-new.row.centers)+center2
  
  new.eigen.vecs <- matrix(0,nrow=n,ncol=n)
  for(j in unique(index)){
    new.eigen.vecs[j,] <- apply(eig.dat2$vectors[which(index==j),,drop=FALSE],2,sum)
  }
  new.coords <- new.centred_mat%*%
    (new.eigen.vecs)%*%diag(eig.dat2$values^(-0.5))

  # Perform Orthogonal Procrustes
  fit.X <- ortho.procrustes(Target=coords[,1:2],Testee = new.coords[,1:2])
  coords2 <- fit.X[[1]]
  
  boot.mds.map.list[[i]] <- coords2
  proc.stat[i] <- fit.X[[2]]
}

boot.mds.map.array <- simplify2array(boot.mds.map.list)
```


### Adding Alpha bags to indicate the sampling variability

```{r, fig.asp=1}

#pdf("plots/palmercRFalphabags.pdf")
a <- par()
par(mar=c(0.1,0.1,0.1,0.1))
alpha <- 0.90
grid.n <- 70

# Plotting spanning ellipses instead of circles
plot(
coords[,1:2],
asp=1,
col=cols,
main="",
xlab="",
ylab="",
pch=16, axes =FALSE, 
frame.plot=TRUE, 
cex=0.2,
type='n'
)

legend(x = "bottomright",legend=levels(class.labels), lty=1, col=col.vec)

a <- sapply(1:dim(boot.mds.map.array)[1],function(x){
  boot.tmp <- t(boot.mds.map.array[x,,])
  boot.tukey.depth <- ddalpha::depth.halfspace(x=boot.tmp,data=boot.tmp, num.directions = 200)
  grid.points <- as.matrix(expand.grid(seq(from=min(boot.tmp[,1]),to=max(boot.tmp[,1]),length.out=grid.n),
                           seq(from=min(boot.tmp[,2]),to=max(boot.tmp[,2]),length.out=grid.n)))
  grid.points.tukey.depth <- ddalpha::depth.halfspace(x=grid.points,data=boot.tmp, num.directions = 200)
  grid.include.ellipse <-which(grid.points.tukey.depth>=quantile(boot.tukey.depth,1-alpha))

  Plot_ConvexHull(xcoord=grid.points[grid.include.ellipse,1],ycoord=grid.points[grid.include.ellipse,2], lcolor=cols[x])
})
#dev.off()
```






## Isomap bootstrap confidence regions based on Tukey depth and alpha bags


```{r}
# n <- nrow(dat)
# #set.seed(123)
# B <- 100
# boot.mat <- matrix(sample(1:n,size=round(n*B,digits = 0),replace=TRUE),nrow=n,ncol=B)
dissim.mat.org <- sqrt(1-extract_proximity_full(tmp.mod,dat))
  s.paths <- stepacross(dissim.mat.org,
                            path='extended',
                            toolong=1, 
                            trace = FALSE)

  dissim.mat <- as.matrix(s.paths)

  A.mat <- -0.5*dissim.mat^2
  col.centers <- colMeans(A.mat)
  row.centers <- colMeans(t(A.mat))
  center <-  mean(A.mat)
  centred_mat <- apply(apply(A.mat,1,function(x)x-col.centers),
                       1,function(x)x-row.centers)+center
  eig.dat <- eigen(centred_mat)
  coords <- eig.dat$vectors[,1:p,drop=FALSE]%*%diag(eig.dat$values[1:p],nrow=p)^0.5

```


```{r, fig.asp=1}
boot.mds.map.list <- list()
proc.stat <- numeric()
for(i in 1:B){
  n <- nrow(tmp.dat)
  index <- boot.mat[,i]
  dist.mat <- dissim.mat.org[index,index]
  s.paths <- stepacross(dist.mat,
                        path='extended',
                        toolong=1, 
                        trace = FALSE)

  dist.mat <- as.matrix(s.paths)
  
  A.mat2 <- -0.5*dist.mat^2
  col.centers2 <- colMeans(A.mat2)
  row.centers2 <- colMeans(t(A.mat2))
  center2 <-  mean(A.mat2)
  centred_mat2 <- apply(apply(A.mat2,1,function(x)x-col.centers2),
                       1,function(x)x-row.centers2)+center2
  
  eig.dat2 <- eigen(centred_mat2)
  
  new.dissim <- dissim.mat
  new.A.mat <- -0.5*new.dissim^2
  new.row.centers <- colMeans(t(new.A.mat))
  new.row.centers[index] <- row.centers2
  new.col.centers <- colMeans(new.A.mat)
  new.col.centers[index] <- col.centers2
  new.centred_mat <- apply(apply(new.A.mat,1,function(x)x-new.col.centers),
                       1,function(x)x-new.row.centers)+center2
  
  new.eigen.vecs <- matrix(0,nrow=n,ncol=n)
  for(j in unique(index)){
    new.eigen.vecs[j,] <- apply(eig.dat2$vectors[which(index==j),,drop=FALSE],2,sum)
  }
  new.coords <- new.centred_mat%*%
    (new.eigen.vecs)%*%diag(eig.dat2$values^(-0.5))

  # Perform Orthogonal Procrustes
  fit.X <- ortho.procrustes(Target=coords[,1:2],Testee = new.coords[,1:2])
  coords2 <- fit.X[[1]]
  
  boot.mds.map.list[[i]] <- coords2
  proc.stat[i] <- fit.X[[2]]
}

boot.mds.map.array <- simplify2array(boot.mds.map.list)

```



```{r, fig.asp=1}
#pdf("plots/palmerisoRFalphabags.pdf")
par(mar=c(0.1,0.1,0.1,0.1))
alpha <- 0.90
grid.n <- 70

# Plotting spanning ellipses instead of circles
plot(
coords[,1:2],
asp=1,
col=cols,
main="",
xlab="",
ylab="",
pch=16, axes =FALSE, frame.plot=TRUE, cex=1
,xlim=range(coords[,1])+c(-0.1,0.05),ylim=range(coords[,2])+c(-0.05,0.1)
,type='n'
)

legend(x = "bottomright",legend=levels(class.labels), lty=1, col=col.vec)

a <- sapply(1:dim(boot.mds.map.array)[1],function(x){
  boot.tmp <- t(boot.mds.map.array[x,,])
  boot.tukey.depth <- ddalpha::depth.halfspace(x=boot.tmp,data=boot.tmp, num.directions = 200)
  grid.points <- as.matrix(expand.grid(seq(from=min(boot.tmp[,1]),to=max(boot.tmp[,1]),length.out=grid.n),
                           seq(from=min(boot.tmp[,2]),to=max(boot.tmp[,2]),length.out=grid.n)))
  grid.points.tukey.depth <- ddalpha::depth.halfspace(x=grid.points,data=boot.tmp, num.directions = 200)
  grid.include.ellipse <-which(grid.points.tukey.depth>=quantile(boot.tukey.depth,1-alpha))

  Plot_ConvexHull(xcoord=grid.points[grid.include.ellipse,1],ycoord=grid.points[grid.include.ellipse,2], lcolor=cols[x])
})

#dev.off()
```


# Monte Carlo Error in Proximity Measure


```{r}
# Extracting the full proximities between every pair of observations
extract_proximity_full_monte_carlo = function(fit, olddata, K=10) {
  pred = predict(fit, olddata, type = "terminalNodes")$predictions
  prox  <- prox.tmp <-  matrix(0, nrow(pred), nrow(pred))
  diag(prox) <- diag(prox.tmp) <- 0.5
  ntree = ncol(pred)
  n = nrow(prox)
  B = ncol(pred)
  
  if (is.null(fit$inbag.counts)) {
    stop("call ranger with keep.inbag = TRUE")
  }
  
  # Get inbag counts
#  inbag = simplify2array(fit$inbag.counts)
  

  for (i in 1:(n-1)) {
    for (j in (i+1):n) {
      prox[i, j] = sum(pred[i, ] == pred[j, ]) / B
    }
  }
  
  
  prox.monte.array <- array(NA, c(n,n,K))
  for (k in 1:K){
    pred.tmp <- pred[,sample(1:B,replace=TRUE)]
    for (i in 1:(n-1)) {
      for (j in (i+1):n) {
        prox.tmp[i, j] = sum(pred.tmp[i, ] == pred.tmp[j, ]) / B
      }
    }
    prox.monte.array[,,k] <- prox.tmp+t(prox.tmp)
  }
  
  
  output <- list(original=prox+t(prox), monte.carlo=prox.monte.array)
  return(output)
  
}
b <- Sys.time()
a <- extract_proximity_full_monte_carlo(mod,tmp.dat,K=2)
bb <- Sys.time()
bb-b
```


```{r, fig.asp=1}
a1.1 <- Sys.time()
ntree <- 50#500
K <- 30
p <- 2
set.seed(12345)
pen.mod.ntree <- ranger(species~.,data=dat,
                  mtry=2,
                  min.node.size = 2, 
                  sample.fraction =0.8264299,
                  num.trees = ntree,
                  keep.inbag=TRUE)
mod <- pen.mod.ntree


prox.monte.carlo.obj <- extract_proximity_full_monte_carlo(mod,tmp.dat,K=K)


dist.mat <- sqrt(1-prox.monte.carlo.obj[[1]])
  A.mat <- -0.5*dist.mat^2
  col.centers <- colMeans(A.mat)
  row.centers <- colMeans(t(A.mat))
  center <-  mean(A.mat)
  centred_mat <- apply(apply(A.mat,1,function(x)x-col.centers),
                       1,function(x)x-row.centers)+center
  eig.dat <- eigen(centred_mat)
  coords <- eig.dat$vectors[,1:p,drop=FALSE]%*%diag(eig.dat$values[1:p],nrow=p)^0.5

prox.mds.map.list <- list()
proc.stat.monte.carlo <- numeric()
prox.monte.carlo.array <- prox.monte.carlo.obj[[2]]
for(k in 1:K){
  dist.mat <- sqrt(1-prox.monte.carlo.array[,,k])
  
  A.mat2 <- -0.5*dist.mat^2
  col.centers2 <- colMeans(A.mat2)
  row.centers2 <- colMeans(t(A.mat2))
  center2 <-  mean(A.mat2)
  centred_mat2 <- apply(apply(A.mat2,1,function(x)x-col.centers2),
                       1,function(x)x-row.centers2)+center2
  
  
  
  eig.dat2 <- eigen(centred_mat2)
  # The coordinates for the factual observations
  coords2 <- eig.dat2$vectors[,1:p,drop=FALSE]%*%diag(eig.dat2$values[1:p],nrow=p)^0.5
  

  # Perform Orthogonal Procrustes
  fit.X <- ortho.procrustes(Target=coords[,1:2],Testee = coords2[,1:2])
  coords2 <- fit.X[[1]]
  
  prox.mds.map.list[[k]] <- coords2
  proc.stat.monte.carlo[k] <- fit.X[[2]]
}

prox.mds.map.array <- simplify2array(prox.mds.map.list)
a1.2 <- Sys.time()
a1.2-a1.1
```

```{r}
mean(unlist(sapply(1:(K-1), function(i) sapply((i+1):K, function(j)sum(diag((prox.monte.carlo.array[,,i]-prox.monte.carlo.array[,,j])%*%t(prox.monte.carlo.array[,,i]-prox.monte.carlo.array[,,j])))))))/(n*(n-1))
summary(proc.stat.monte.carlo)
```

```{r}
#Analytical Estimate
a <- (prox.monte.carlo.obj$original[upper.tri(prox.monte.carlo.obj$original,diag = FALSE)])
2*sum(a*(1-a))/length(a)/ntree

```


```{r, fig.asp=1}
for(i in 1:5){
  plot(prox.mds.map.array[,,i],col=cols)
}
```


```{r,fig.asp=1}
#pdf("plots/palmerConteCarloBootRFs_ntree.pdf")
par(mar=c(0.1,0.1,0.1,0.1))
alpha <- 1

# Plotting spanning ellipses instead of circles
plot(
coords[,1:2],
asp=1,
col=cols,
main="",
xlab="",
ylab="",
pch=16, 
axes =FALSE, 
frame.plot=TRUE,
type="n"
)


legend(x = "bottomright",legend=levels(Class), pch=16, col=col.vec)

a <- sapply(1:dim(prox.mds.map.array)[1],function(x){
  prox.tmp <- t(prox.mds.map.array[x,,])
  Plot_ConvexHull(xcoord=prox.tmp[,1],ycoord=prox.tmp[,2], lcolor=cols[x])
})
#dev.off()
```


### Prox Plot Monte Carlo with new RFs for every plot


```{r, fig.asp=1}
a2.1 <- Sys.time()
set.seed(12345)
ntree <- 500
K <- 50
p <- 2
n <- nrow(tmp.dat)
pen.mod.ntree <- ranger(species~.,data=dat,
                  mtry=2,
                  min.node.size = 2, 
                  sample.fraction =0.8264299,
                  num.trees = ntree,
                  keep.inbag=TRUE)
mod <- pen.mod.ntree


prox.monte.carlo.obj <- extract_proximity_full(mod,tmp.dat)
prox.array <- array(NA,dim=c(n,n,K))

dist.mat <- sqrt(1-prox.monte.carlo.obj)
  A.mat <- -0.5*dist.mat^2
  col.centers <- colMeans(A.mat)
  row.centers <- colMeans(t(A.mat))
  center <-  mean(A.mat)
  centred_mat <- apply(apply(A.mat,1,function(x)x-col.centers),
                       1,function(x)x-row.centers)+center
  eig.dat <- eigen(centred_mat)
  coords <- eig.dat$vectors[,1:p,drop=FALSE]%*%diag(eig.dat$values[1:p],nrow=p)^0.5

prox.mds.map.list <- list()
proc.stat <- numeric()
for(k in 1:K){
  pen.tmp.mod.100 <- ranger(species~.,data=dat,
                  mtry=2,
                  min.node.size = 2, 
                  sample.fraction =0.8264299,
                  num.trees = ntree,
                  keep.inbag=TRUE)
  pen.prox.mat <- extract_proximity_full(pen.tmp.mod.100,tmp.dat)
  prox.array[,,k] <- pen.prox.mat
  dist.mat <- sqrt(1-pen.prox.mat)
  
  A.mat2 <- -0.5*dist.mat^2
  col.centers2 <- colMeans(A.mat2)
  row.centers2 <- colMeans(t(A.mat2))
  center2 <-  mean(A.mat2)
  centred_mat2 <- apply(apply(A.mat2,1,function(x)x-col.centers2),
                       1,function(x)x-row.centers2)+center2
  
  
  
  eig.dat2 <- eigen(centred_mat2)
  # The coordinates for the factual observations
  coords2 <- eig.dat2$vectors[,1:p,drop=FALSE]%*%diag(eig.dat2$values[1:p],nrow=p)^0.5
  

  # Perform Orthogonal Procrustes
  fit.X <- ortho.procrustes(Target=coords[,1:2],Testee = coords2[,1:2])
  coords2 <- fit.X[[1]]
  
  prox.mds.map.list[[k]] <- coords2
  proc.stat[k] <- fit.X[[2]]
}
prox.mds.map.array <- simplify2array(prox.mds.map.list)
a2.2 <- Sys.time()
a2.2-a2.1
```

```{r}
mean(unlist(sapply(1:(K-1), function(i) sapply((i+1):K, function(j)sum(diag((prox.array[,,i]-prox.array[,,j])%*%t(prox.array[,,i]-prox.array[,,j])))))))/(n*(n-1))
summary(proc.stat)
```


```{r, fig.asp=1}
for(i in 1:K){
  plot(prox.mds.map.array[,,i],col=cols)
}
```


```{r,fig.asp=1}
#pdf("plots/palmerConteCarloNewRFs_ntree.pdf")
par(mar=c(0.1,0.1,0.1,0.1))
alpha <- 1
grid.n <- 50

# Plotting spanning ellipses instead of circles
plot(
coords[,1:2],
asp=1,
col=cols,
main="",
xlab="",
ylab="",
pch=16, 
axes =FALSE, 
frame.plot=TRUE,
type="n"
)


legend(x = "bottomright",legend=levels(Class), pch=16, col=col.vec)



a <- sapply(1:dim(prox.mds.map.array)[1],function(x){
  prox.tmp <- t(prox.mds.map.array[x,,])
  Plot_ConvexHull(xcoord=prox.tmp[,1],ycoord=prox.tmp[,2], lcolor=cols[x])
})
#dev.off()
```











